---
title: "Task1"
author: "Anton Stadnik"
date: "14/10/2025"
format:
  html:
    embed-resources: true
editor: source
---

## Question 1 title

Question description

```{r}
# code for solving task
```


## Question N title

Question description

```{r}
# code for solving task

# =========================================================
# Task 1: Exploring Data Types in R
# Theme: Lvyv — The Floating Sky City
# =========================================================

# -----------------------------
# 1.1 Atomic Types
# -----------------------------

# numeric values
avg_temperature <- 18.7     # degrees Celsius
altitude_km     <- 7.5      # floating height above ground (km)
energy_rating   <- 92.4     # sustainability rating
water_reserves  <- 35.2     # in million liters

# integer values
population      <- as.integer(12000)
watchtowers     <- as.integer(8)
airships        <- as.integer(42)

# logical values
has_magic_crystals <- TRUE
is_allied_empire   <- FALSE

# character values
city_name   <- "Lvyv"
region_name <- "Sky Realm"
city_type   <- "Floating Metropolis"

cat("The city of", city_name, "is a", city_type, "in the", region_name,
    "floating at", altitude_km, "km with a population of", population, "\n")

# -----------------------------
# 1.2 Vectors
# -----------------------------

numeric_vector <- c(avg_temperature, altitude_km, energy_rating, water_reserves,
                    population, watchtowers, airships)

sequence_vector <- seq(20, 100, by = 5)

set.seed(123)
logical_vector <- sample(c(TRUE, FALSE), size = 10, replace = TRUE)

# -----------------------------
# 1.3 Matrix
# -----------------------------

set.seed(42)
lvyv_matrix <- matrix(sample(-50:50, 9, replace = TRUE),
                      nrow = 3, ncol = 3)

rownames(lvyv_matrix) <- c("Tower", "Docks", "Garden")
colnames(lvyv_matrix) <- c("A", "B", "C")

row_means <- rowMeans(lvyv_matrix)
col_means <- colMeans(lvyv_matrix)

lvyv_matrix
row_means
col_means

# -----------------------------
# 1.4 List
# -----------------------------

lvyv_list <- list(
  city_name = city_name,
  numeric_vector = numeric_vector,
  sequence_vector = sequence_vector,
  logical_vector = logical_vector,
  infrastructure_matrix = lvyv_matrix,
  description = "Lvyv is a floating city sustained by magic crystals and airships, hovering gracefully above the clouds."
)

lvyv_list

# -----------------------------
# 1.5 Factor
# -----------------------------

econ_levels <- factor(c("high", "medium", "medium", "high", "low", "medium", "high"))
table(econ_levels)

# -----------------------------
# 1.6 Data Frame
# -----------------------------

lvyv_df <- data.frame(
  sector = c("Central Spire", "Harbor Deck", "Crystal Forge", "Sky Garden",
             "Guild Quarters", "Aviary"),
  visitors_per_day = c(500, 350, 200, 450, 180, 250),
  is_restricted = c(TRUE, FALSE, TRUE, FALSE, FALSE, TRUE),
  econ_class = factor(c("high", "medium", "high", "medium", "low", "medium"))
)

colnames(lvyv_df) <- c("Sector", "Visitors", "Restricted", "EconClass")

str(lvyv_df)
summary(lvyv_df)

lvyv_df$TrafficLevel <- ifelse(lvyv_df$Visitors > 300, "high", "low")

lvyv_df

